Evolution of Sex-Specific Recombination Rate

### 2020-05-08 doc/MS_v1.Rmd
fixing typos and updating the draft

We used the single crossover bivalent (1CO) normalized position to compared general location of crossovers while controlling for differences in total SC length. Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.

@@

Regardless of whether these hypotheses are true, our observations of between cell variation fail to support a basic prediction of (the relationship between mean and variance) a higher mean will be associated with a higher variance. Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count. These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002, @gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes. This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].

In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure. This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals. Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007]. Centrosomes spindles are in spermatogenesis and mature sperm cells delivers centrioles (most mammals) or centrosome material (rodents and rabbits) to the egg during fertilization [@manandhar2005centrosome]. This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].

Regardless of whether these hypotheses are true, our results further support the results of females having greater variation in the number of crossovers across cells in mice and humans [@lynn2002, @gruhn2013]. This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].

>>>>

We used the single crossover bivalent (1CO) position to compared general location of crossovers while controlling for differences in total SC length. Given that, in most of the strains we surveyed, the majority of bivalents are observed to contain one crossover (focus), thus justifying the focus of single crossover bivalents for analyzing crossover positions.

@@

In the spindle based selection model, the less efficient SAC in oocytes leads to relaxed selection on the metaphase I chromosome structure. This will result in increased variance in chromosome structure and crossover number per cell, not just across strains, but across cells within individuals. Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers [@schuh2007]. Centrosomes act as the two MTOCs 
for the spindle during spermatogenesis and mitosis. The delivery of centrioles (most mammals) or centriole material (rodents and rabbits) to the egg is thought to be a barrier to evolution of parthenogenesis [@manandhar2005centrosome].This is another way the spindle based selection model is linked to conserved sex differences in gametogenesis [@ross2015].

Regardless of whether these hypotheses are true, our observations of levels of between cell variation fail to support a basic prediction of that a higher mean will be associated with a higher variance. Across all strains, including the high-recombining strains, _muscuslus^PWD^_, _muscuslus^SKIVE^_ and _molossinus^MSM^_, females consistently have higher between cell variance for crossover count. These results add to previous findings in mice and humans that oogenesis has increased variance for crossovers [@lynn2002], [@gruhn2013] and precursors to crossovers [@lenzi2005], compared to spermatocytes. This feature of heterochiasmy supports evolutionary theories predicting distinct reproductive strategies between females (diversifying offspring) and males (maintain successful haplotypes) [@trivers2002].
